---
title: "An Update on Broom"
author: "Alex Hayes"
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    keep_tex: no
    theme: metropolis
    slide_level: 2
    incremental: no
fontsize: 12pt
classoption: compress
header-includes:
  \setbeamercolor{frametitle}{bg=gray}
  \hypersetup{colorlinks,citecolor=orange,filecolor=red,linkcolor=brown,urlcolor=blue}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(broom)
```

## Outline

1. Progress so far
    - Bug fixes and pull requests
    - New test suite
    - New documentation
2. broom 0.5.0 release
3. Lessons learned
4. The future of broom

# What I've been working on

## Bug fixes and pull requests

- Took a while to clean up the issues
- PRs are of highly varying quality
    - Tests often missing or limited
    - Documentation often sparse and not particularly useful
- Bug fixes generally mellow (tidiers for a given class rarely more than 300 lines, including documentation)
- High levels of enthusiasm

## Bug fixes and pull requests: adding tidiers vignette

Work in progress
Describes standards for contributing

https://broom.tidyverse.org/articles/adding-tidiers.html

## Test suite: coverage

Before 0.5.0, line coverage was pretty bad (~40 percent)

Now all tidying methods have at least one test

Coverage at 80 percent (higher in practice -- long running tests for deprecated tidiers skipped)

Example: `tidy.Arima()` has "high coverage"

Line coverage less important than model coverage: what variety of models with different options can we tidy

```{r, eval = FALSE}
glance(arima(lh, order = c(3,0,0)))
glance(arima(lh, order = c(3,0,0), method = "CSS"))  # broke until very recently
```

Need coverage in the model object space

PRs often come without tests, or with tests only for the subset of the model object space that the PR author uses in their day to day

If you don't know the model, it's hard to write this sort of stress test

## Test suite: tibble output

- `tidy()`, `glance()`, `augment()` should return tibbles
- `glance()` should return a single row
- `augment()` tests still in progress, trickier than expected

```{r}
fit <- lm(hp ~ ., mtcars)
td <- tidy(fit)
check_tidy_output(td)
```

## Test suite: argument checking

```{r}
check_argument(tidy.lm)
```

- Check arguments against master list
- Check for default arguments
    - lots of `is.missing()` in old tidiers
    - `conf.int = FALSE`
    - `conf.level = 0.95`
    - `conf.int` and `conf.level` always come as a pair
    
Goal: enforce consistency, especially in new PRs

Did this manually for 0.5.0, will be automated in 0.7.0

## Test suite: column naming

```{r, eval = TRUE}
library(lavaan)

cfa.fit <- cfa(
   'F =~ x1 + x2 + x3 + x4 + x5',
   data = HolzingerSwineford1939, group = "school"
)

glance(cfa.fit)
```

These names are inconsistent with the rest of broom

Goal: again, put consistency burden onto PR authors

## Test suite: column naming strategy

For 0.7.0:

- Describe acceptable column names in `tidy.yaml`. Excerpt:

```{yaml}
- column: adj.r.squared
  description: Adjusted R squared statistic, taking into account degrees of freedom.
  used_by:
    - ivreg
```

- Compile this into a `column_glossary` tibble
- Export `column_glossary` (downstream package maintainers have asked for this)
- Test output column names against `column_glossary`
- Use `roxygen2` templates to populate documentation from `column_glossary`

## Documentation: templates

Many repeated arguments:

```{r, eval = FALSE}
tidy.betareg <- function(x, conf.int = FALSE, conf.level = .95, ...) 
tidy.ivreg <- function(x, conf.int = FALSE, conf.level = .95, exponentiate = FALSE, ...)
```

All `conf.int` argument behave the same way: should have the same documentation.

## Documentation: templates

Now `man-roxygen/param_confint.R` does this:

```
#' @param conf.int Logical indicating whether or not to include a confidence
#'   interval in the tidied output. Defaults to `FALSE`.
#' @param conf.level The confidence level to use for the confidence interval
#'   if `conf.int = TRUE`. Must be strictly greater than 0 and less than 1.
#'   Defaults to 0.95, which corresponds to a 95 percent confidence interval.
#' @md
```

Drop into documentation with

```
@template param_confint
```

## Documentation: templates

Templates instead of inheriting parameters because:

- I learned about templates first
- Couldn't figure out how to inherit doc for only a single parameter
- Can run arbitrary R code inside them

Templates currently used for:

- Titles
- Descriptions
- Parameters
- Some standardized types of returns

## Documentation: individualize documentation

Old strategy:

- document `tidy.lm()`, `glance.lm()` and `augment.lm()` all in `lm_tidiers.Rd`
- stuff like this:

```{r, eval = FALSE}
#' @rdname tidy.lm
#' @export
tidy.rlm <- tidy.lm
```

Now every function has it's own documentation and Rd file

- Less magical
- Documentation easier to find
    - `tidy.glm()` didn't have documentation until two weeks ago
- Heavily cross-linked and aliased

## Documentation: return columns

Goal for 0.7.0: populate `#' @return` from `column_glossary`.

# broom 0.5.0

## broom 0.5.0: features

- Tibble output
- Test suite
- Documentation
- ~10 new tidiers (all contributed)
- Tons of bug fixes (mostly contributed, some by me)

## broom 0.5.0: tibble output

Tibbles break some things, mostly when:

- subsetting with `[` and expecting a vector.
- setting rownames on a tibble.
- using augment on models making use of matrix covariates / outcomes.
    - i.e. `survival::Surv()`
    
## broom 0.5.0: matrix column and augment example

Data with matrix-column:

```{r, eval = TRUE}
y <- rnorm(5)
x <- matrix(rnorm(10), nrow = 5)
df <- data.frame(x, y)
```

Tibbles don't like this:

```{r, eval = TRUE, error = TRUE}
tibble::tibble(x, y)
```

Modelling functions don't mind:

```{r, error = TRUE}
fit <- lm(y ~ x, df)
augment(fit)
```

Sometimes passing `data` argument helps:

```{r}
augment(fit, data = df)
```

## broom 0.5.0: deprecations

- Broom tidiers some non-statistical objects
- Moving away from this. Deprecating
    - `tidy.data.frame()`
    - `tidy.matrix()`
    - `tidy.numeric()`
    - `tidy.vector()`
- Should use `tibble::as_tibble()` instead
- Couple more of these coming in 0.7.0

## broom 0.5.0: deprecations: mixed models

Moving tidiers for 

- lme, lme4 and nmle models,
- brms models,
- rstanarm models, and 
- mcmc objects

to Ben Bolker's `broom.mixed` package

# Lessons learned

## Making systematic changes is time consuming

- 100+ tidiers
- Model objects are simple but often unfamiliar
- Changing all tidiers (i.e. new tests/doc) take 1-1.5 weeks

## Broom depends on high quality PRs

Very dependent on community members to get reasonable tidiers -- figuring out what should get tidied is not a straightforward process if you know nothing about the model

Writing a low quality PR can take more time than writing one from scratch -- trying to improve this situation via documentation and testing tools

## People sometimes use broom to do weird things



In the process of updating to tibbles:
- lots of people still set rownames on dataframes

From deprecations:
- many people are using broom to coerce to tibble

## `augment()` is hard

Original thought: `tidy()` is most ambiguous method, will be hardest to work with

Incorrect: `augment()` is hard

- Need different behavior for `data` and `newdata` args
- People often don't implement it
- Have to deal with model both model input and output

# The Future of Broom

## The big split

What: split broom into domain specific tidying packages
Why: high maintanence and design burden

**Delays**:

- Internals are messier than anticipated
- Tidier behavior not fully specified
    - `augment()` NA behavior
- Lots of tidiers don't meet existing specifications

## The big split: vision

- import `tidy()`, `glance()` and `augment()` from `modelgenerics`
- `broom` tidies models in `base` and `stats`
    - maybe some other popular models
    - smaller, fewer dependencies
- broom documentation defines tidier specifications
- domain-specific packages import broom
- broom + domain specific package live in `tidymodels`
- some system for tracking where tidiers live

```{r}
library(tidymodels)  # load everything
```

## Possible domain specific packages

- sweep (on CRAN, mature)
- tidytext (on CRAN, mature)
- broomstick (recently started work in progress)
- broom.mixed (not on CRAN, at about the same level of development as broom)
- biobroom (on BioConductor, mature)
- broom.base (work in progress for low dependency package to act as base for extensions)
- Gavin Simpson (schoenberg??)
- tidybayes (mature, preparing for CRAN release)
    - operates on posterior samples, not model objects

## Timeline

TBD pending meeting with Dave tomorrow

- July: Finalize tidier specifications
- July: Start collaborating on domain specific packages
- Early August: Implement as much spec as possible for 0.7.0 release
- August 20: internship ends
- Late August: Release broom 0.7.0
    - I will like become package maintainer
- Early September: grad school starts (dev slows down)
- September+: rewrite the broom paper with Dave
- September+

# Questions?

LINKS LINKS LINKS

@alexpghayes on Twitter  
alexpghayes@gmail.com

